# Datasets to Practice

## Environment Setup

We use the tidyverse suite for data manipulation and visualization.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(scales) # For professional axis labels
```

### The Tidyverse Collection (Modern Data Science)
The `tidyverse` is a suite of packages designed for a seamless data workflow. Its datasets are primarily used to master **Data Wrangling**, **Tidy Data principles**, and **Advanced Visualization**.

| Dataset | Package | Description | Primary Analysis Use |
| :--- | :--- | :--- | :--- |
| **`mpg`** | `ggplot2` | Fuel economy data for 38 car models (1999–2008). | **Exploratory Data Analysis (EDA)**: Practicing scatterplots, smoothing lines, and correlation between engine size and efficiency. |
| **`diamonds`** | `ggplot2` | Prices and attributes of ~54,000 diamonds. | **Predictive Modeling**: Ideal for training linear regression models or random forests due to its large sample size. |
| **`starwars`** | `dplyr` | Characteristics of characters (height, mass, species). | **Data Cleaning**: Best for practicing how to handle `NA` (missing values) and complex list-columns. |
| **`table1-5`** | `tidyr` | WHO Tuberculosis cases in different formats. | **Data Reshaping**: Specifically used to learn `pivot_longer()` and `pivot_wider()` to reach a "tidy" state. |
| **`storms`** | `dplyr` | Wind and pressure data for tropical cyclones. | **Time Series & Mapping**: Visualizing storm tracks over time and geographical coordinates. |

---

### ISwR (Introductory Statistics with R)
The `ISwR` package (by Peter Dalgaard) focuses on **Biostatistics** and **Inference**. These datasets are smaller but are "textbook cases" for specific mathematical tests.

| Dataset | Description | Primary Analysis Use |
| :--- | :--- | :--- |
| **`thuesen`** | Ventricular shortening velocity and blood glucose in diabetics. | **Simple Linear Regression**: Testing the linear relationship between two continuous clinical variables. |
| **`energy`** | Energy expenditure in lean vs. obese women. | **Hypothesis Testing**: Perfect for practicing the Two-sample t-test and checking for equal variance. |
| **`juul`** | IGF-I (Growth Hormone) levels across different ages. | **ANOVA & Polynomial Regression**: Analyzing how a biological marker changes non-linearly across puberty stages. |
| **`melanom`** | Survival data of patients with malignant melanoma. | **Survival Analysis**: The gold standard for practicing Kaplan-Meier survival curves and Cox Proportional Hazards. |
| **`stroke`** | Clinical outcomes of stroke patients in Estonia. | **Logistic Regression**: Predicting binary outcomes (e.g., dead/alive) based on age, sex, and clinical history. |

---


## Loading the dataset

```{r}
data("diamonds")
```
<!-- 
3. Exploratory Data Analysis (EDA)

Before modeling, we must understand the distribution of our target variable: price.

3.1 Distribution of Price

Most diamonds are relatively inexpensive, but there is a "long tail" of high-value stones.

diamonds |> 
  ggplot(aes(x = price)) +
  geom_histogram(bins = 50, fill = "#2c3e50", color = "white") +
  scale_x_continuous(labels = label_dollar()) +
  labs(
    title = "Distribution of Diamond Prices",
    x = "Price (USD)",
    y = "Frequency"
  ) +
  theme_minimal()


3.2 The Carat-Price Relationship

The relationship between weight (carat) and price is clearly non-linear. As carat increases, the price increases exponentially—a classic example of "diminishing returns" in reverse (increasing returns to rarity).

diamonds |> 
  slice_sample(n = 5000) |> # Sampling for clearer visualization
  ggplot(aes(x = carat, y = price)) +
  geom_point(alpha = 0.2, color = "#3498db") +
  geom_smooth(method = "gam", color = "red") +
  scale_y_continuous(labels = label_dollar()) +
  labs(
    title = "Price vs. Carat Weight",
    subtitle = "Non-linear relationship suggests a power-law or exponential trend",
    x = "Weight (Carat)",
    y = "Price (USD)"
  ) +
  theme_minimal()


4. Feature Engineering: Log Transformations

In industry, we often transform skewed data to meet the assumptions of linear models. By applying a log10 transformation to both Price and Carat, we can linearize the relationship.

# Linearizing the relationship
diamonds |> 
  ggplot(aes(x = log10(carat), y = log10(price))) +
  geom_point(alpha = 0.1, color = "#2ecc71") +
  geom_smooth(method = "lm", color = "black") +
  labs(
    title = "Log-Log Transformation: Price vs. Carat",
    subtitle = "A clear linear trend emerges after transformation",
    x = "Log10(Carat)",
    y = "Log10(Price)"
  ) +
  theme_minimal()


5. Statistical Modeling

We will fit a linear model using the log-transformed variables to predict price.

# Fitting the model
model <- lm(log10(price) ~ log10(carat) + cut + color + clarity, data = diamonds)

# Summary of the model
summary(model)


Key Insights:

R-squared: The high R-squared value indicates that the "Four Cs" explain nearly 98% of the variance in (logged) prices.

Carat Impact: For every 1% increase in carat weight, the price increases by approximately 1.6% (interpreting the log-log coefficient).

6. Conclusion

This analysis demonstrates that while Carat is the primary driver of price, the categorical variables (Cut, Color, Clarity) act as significant multipliers. For a professional portfolio, this workflow shows a transition from raw data exploration to a mathematically sound predictive framework.

Note: This notebook was created for self-study and portfolio purposes using R 4.4.1. -->