# Hypothesis Test

## mean inference (z,t-test)

### Z-test

Note that there's no built-in Z-test in R. Use third party package ```BSDA``` (the code above written by Gemini)

```{r}
# install.packages("BSDA")
library(BSDA)

x <- c(1,1,1,1,1,4,3,2,1,1,1,1,1,11,1,1)
z.test(x, sigma.x = 0.5, conf.level = 0.95)
```

You can also make implement from the formular: reference this passage: https://cran.r-project.org/web/packages/distributions3/vignettes/one-sample-z-test.html

### t-test

```{r}
x <- c(1,1,1,1,1,4,3,2,1,1,1,1,1,11,1,1)
y <- c(1,1,1,1,1,4,3,2,1,1,1,1,1,11,1,1)
t.test(x)
t.test(x,y)
t.test(x,y, mu = 2, alter = "two.sided", var.equal = T)

```


## variance inference

Try to type ```?var.test``` to get the help document.
```{r}
var.test(x,y)

```


## goodness of fitness test

Test whether a sample data fit a distribution or not.

### Normal Test

## correlation

```{r}
x <- c(1,1,1,1,1,4,3,2,1,1,1,1,1,11,1,1)
y <- c(1,1,1,1,1,4,3,2,1,1,1,1,1,11,1,1)
cor.test(x, y)

```

The default method, Pearson linear correlation is widely used. 

## median inference (non-para test)

Although it's non-parametric test based on median, the function also require to pass parameter as ```mu```.

```{r}

x <- c(1,1,1,1,1,4,3,2,1,1,1,1,1,11,1,1)
y <- c(1,1,1,1,1,4,3,2,1,1,1,1,1,11,1,1)
wilcox.test(x)
wilcox.test(x,y)
wilcox.test(x,y, mu = 2, alter = "two.sided", var.equal = T)

```


