# Monte Carlo

```{r, echo = FALSE}
set.seed(114514)
```
## Monte Carlo Simulation of Intergation

### Example 1
Estimating this intergal by Monte Carlo:
$$
\theta = \int_{0}^{1} {e^x} dx = e-1
$$

The real integation is $e-1$. But this is meaningless for those we can't integrate easily.

The idea of Monte Carlo is, assuming there's a area x from 0 to 1, y from 0 to M (a large number, greater than the maximum of function we're estimating)

```{r, echo = FALSE}
curve(exp(x), from = 0, to = 1, xlim = c(-0.5, 1.5), ylim = c(-0.5, 4))
lines(c(-0.5, 1.5), c(0,0))
lines(c(0, 1), c(3,3))
lines(c(0,0), c(-0.5, 4))
lines(c(1,1), c(0, 4))

x_shade <- seq(0, 1, 0.01)
y_shade <- exp(x_shade)

polygon(c(0, x_shade, 1), c(0, y_shade, 0), 
        col = rgb(0.1, 0.6, 0.1, alpha = 0.3), border = NA)

text(0.5, 1.2, labels = expression(e^x), cex = 1.5, col = "darkblue")
text(1.1, exp(1), labels = "e", pos = 4)
text(0.5, 0.5, labels = "Area", font = 2)
```

We're calculting the green area in the plot, note that there's a probability of random point in the green area is equal to their ratio of area.

Find any M, i.e., here we let $M = 3 > e$

$$
p = \frac{\text{Area of Green}}{\text{Area of Rectangular}} = \frac{S_G}{S_R}
$$

Let $X$ is a random variable with uniform distribution from 0 to 1,  $Y = e^X$, then the Intergal $\hat{\theta} = S_G = p S_R = \mathbb{E}[\bar{Y}]$


```{r}
m <- 100000
x <- runif(m)
c("Monte Carlo", "Real Intergation")
c(mean(exp(x)), exp(1)-1)
```

### Example 2: Improper Integral


Estimating this intergal by Monte Carlo: (F(x) here is CDF of normal distribution)
$$
\theta = \int_{-\infty}^{x} {e^{-\frac{t^2}{2}}} dt  = \sqrt{2\pi} F(x)
$$


<!-- = \frac{1}{2} + \frac{\sqrt{\pi}}{2} erf(x) -->

Let's make transformation first:

$$\theta = \sqrt{2\pi} (0.5+ \int_{0}^{x} {e^{-\frac{t^2}{2}}} dt )=\sqrt{\frac{\pi}{2}}+ \sqrt{2\pi} \int_{0}^{1} xe^{\frac{(xy)^2}{2}}dy$$

```{r}
m <- 100000
y <- runif(m)
x <- 1
c("Monte Carlo", "Real Intergation")
c(sqrt(pi)/2+ x * mean(exp((x*y)^2/2)),sqrt(pi*2) * pnorm(1))
```

Another idea is that, the sample EDF will converge to CDF. Thus, let's generate normal random variables.

```{r}
m <- 100000
y <- rnorm(m)
x <- 1

c(sqrt(pi*2) * mean(y<=x) ,sqrt(pi*2) * pnorm(1))
```


### Variance


In the first example we use $Y = e^X$ to estimate $ \theta = \int_{0}^{1} {e^x} dx $

As $X\sim U(0,1)$, it's unbiased estimation since $\mathbb{E}[Y] = \mathbb{E}[e^X] = e-1 = \theta$.

The variance is $Var[Y] = \mathbb{E}[Y^2] - {[\mathbb{E}[Y]]}^2 = \frac{(e-1)(3-e)}{2} = 0.2420$

```{r, echo = F}
print(paste(c("variance is, ",(exp(1)-1)*(3-exp(1))/2),sep = "", collapse= ""))
```

For the second example, we have two estimation: 

$$X_1 \sim U(0,1) \Rightarrow Y_1 = \frac{\sqrt{\pi}}{2} + x e^{\frac{xX_1}{2}}$$


```{r}
m <- 100000
x_1 <- runif(m)
x <- 1
y_1 <- sqrt(pi)/2+ x * exp((x*x_1)^2/2)
var(y_1)
```

$$X_2 \sim N(0,1) \Rightarrow \frac{Y_2}{\sqrt{2\pi}}\sim B(1,p)$$


```{r}
m <- 100000
x_2 <- rnorm(m)
x <- 1
y_2 <- sqrt(pi*2) * (y<=x)

var(y_2)
```
