# R Basics

## Generate Random Sample

set random seed to get same result for repeating
```{r}
set.seed(42)
```

a random vector size of 20, follows to Normal Distribution
```{r}
rnorm(20)
rnorm(20, mean = 10, sd = 2)

```

a random vector size of 10, follows to Uniform Distribution

```{r}
runif(20)
runif(20, min = 0, max = 1)
runif(20, 0, 1)

```


```{r}
prod(40:36)
prod(5:1)/prod(40:36)
1/choose(40,5)
```

```{r}
pvec <- seq(0,1,0.05)
pvec

x <- rnorm(60000)
quantile(x, pvec)
quantile(x)
```


```{r}
x.1 <- rnorm(20)
x.2 <- rnorm(20) + runif(20)
x.3 <- x.1 * 20 + x.2 * 4 - rnorm(20)

df <- data.frame(x1 = x.1, x2 = x.2, x3 = x.3)
```

## Matrix
```{r}
A <- cor(df)
A


svd(A)
iA <- solve(A)

cA <- chol(A)

t(cA)  %*% cA


solve(cA) %*% t(solve(cA))
solve(A)
```



## string
```{r}
astr <- as.character(12345)
bstr <- "NULL"
cstr <- "NOT LOWER, not upper"
c(astr, bstr, cstr)

```



### length of a string
```{r}
nchar(astr)
```


```{r}
c(toupper(cstr), tolower(cstr))

```



### substring

both ```substr``` and ```substring``` can extract substring, and ```substring``` defined with default value for the ```last = 1000000L```. Also, ```substring``` can exact multiple substrings while ```substr``` can't.
```{r}
substr("HELLO", 2,3)
substr("HELLO", c(1,2,3),c(3,4,5))
substring("HELLO", c(1,2,3),c(3,4,5))
```



```{r}
a = "12345678"
substr(a,2,2)
substr(a,2,4)

# substr(a,2)
# Error in substr(a, 2) : argument "stop" is missing, with no default

substring(a,2)
substr(a,3,4) <- "Toyota"
a

substring(a,3) <- "Toyata"
a
```



### combine strings
```{r}

a = "12345"

b = "6789"

paste(a,b)
paste(a,b,sep="")
paste(a,b,sep="-")
paste(a,b,sep = "toyota")

a <- c('Taipei', "Tokyo", 'Singapore', 'Hong Kong', "Johor Bahru", "Kuala Lumpur")
paste(a)
paste(a, "?", sep = "")
paste(a, "is where", sep = " ")
paste(a, "is where", sep = " ", collapse = ", and ")


paste(c("a","b"), c("1","2","3"))
paste(c("a","b","c","d"), c("1","2","3"))

```

### split strings
```{r}
strsplit(",,a,bb,", split = ",")
strsplit(",,a,,", split = ",")
strsplit(",,a,", split = ",")
```


### replace

replace by vector
```{r}
s <- "ccccaaaassssaaa"
chartr("s", "A", s)
chartr("sc", "AI", s)
sub("a", "b", s)
gsub("a", "b", s)

```

Note that ```chartr``` is vector-behavior function, which changes char-by-char; as ```gsub``` replaces the whole target substring globally, ```sub``` only changes the first occurance. 
### Regular Expression 
try: type ```?grep``` in your console. Then see the help page "Pattern Matching and Replacement"

```{r}
# grep()
```

## todo

```{r}
c("Hello")

```


```{r}
c("Hello")

```


```{r}
c("Hello")

```


```{r}
c("Hello")

```


```{r}
c("Hello")

```


```{r}
c("Hello")

```


```{r}
c("Hello")

```


```{r}
c("Hello")

```


```{r}
c("Hello")

```


```{r}
c("Hello")

```


```{r}
c("Hello")

```


```{r}
c("Hello")

```


```{r}
c("Hello")

```


```{r}
c("Hello")

```


```{r}
c("Hello")

```

## tricks
### deal with potential exception
pass a string of the variable name to check if it's defined, return NULL if not found.
```{r}
x0 <- "demo"
get0("x0")
get0("xa_nondefined")

```

### RMarkdown features

When you write {r} after three ` signs you can give some features: 

You can leave {r} by default and get very annoying outputs
```{r}
library(dplyr)
library(tidyverse)
```

you can write ```{r, warning=FALSE, message=FALSE}``` means no suppresses warnings and no suppresses messages.

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyverse)
```

```echo = FALSE``` means to hide the code trunk

```include=FALSE``` means to hide the code trunk and the result

```{r, echo = FALSE}
library(dplyr)
library(tidyverse)
```


```eval = FALSE``` show but no executing


```results = 'hide'``` will hide the code 


```error = TRUE``` continue to knit even there's error


### package management
install if not exists then load the package.
```{r, warning=FALSE, message=FALSE}
if (!requireNamespace("tidyverse")) install.packages('tidyverse')
library(tidyverse)
```


This can also be made with package ```pacman```.
```{r, warning=FALSE, message=FALSE}
library(pacman)
pacman::p_load(mblm)

```


```{r}
c("Hello")

```